---
title: "RSA analysis of Project DRIVE data"
author: "John Bruce"
date: "11/13/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This is the code to process the DRIVE raw data to feed it into the RSA algorithm.

Data was obtained from the Mendeley Web site (https://data.mendeley.com/datasets/y3ds55n88r/4). Two data files are used: DriveCountData.RDS and ShrnaGeneMap.RDS

The file DriveCountData has data for three pools, poolA, poolB, and BGPD. The data can be analyzed all together, as well as for each pool separately.

Below is the processing of all data together. Farther down, the separate processing for the three pools is shown.

```{r}
setwd('/data')
con <- gzfile('DriveCountData.RDS','rb')
countdata <- readRDS(con)
countdata <- subset(countdata, !is.na(PLASMID_COUNT), !is.na(SAMPLE_COUNT) )
library(edgeR)
y.all <- DGEList(counts=countdata[,3:4],genes=countdata[,c(1,2,5,6,7)])
num_samples <- dim(countdata)[1]
total_plasmid_count <- y.all$samples[1,2]
avg_plasmid_count <- round(total_plasmid_count / num_samples)
```
We impute plasmid count for elements of PLASMID_COUNT column that are NA, with avg_plasmid_count 
(they were made NA because they had low values)
```{r}
countdata[is.na(countdata[,3]),3] <- avg_plasmid_count
y.all <- DGEList(counts=countdata[,3:4],genes=countdata[,c(1,2,5,6,7)])
y.all <- calcNormFactors(y.all)
treat <- factor(substring(colnames(countdata)[3:4],1,5))
treat <- relevel(treat, ref="PLASM")
design <- model.matrix(~treat)
```
The edgeR negative binomial model is fitted to obtain log fold change (logFC) of counts
between the sample and the plasmid. This is done for each sample-plasmid pair, to obtain
a logFC per shRNA per cell line.
```{r}
fit <- glmFit(y.all, design, dispersion=0.2, prior.count=12)
fit2 <- glmLRT(fit)
head(fit2$table)
countdata <- cbind(countdata,fit2$table)
# saved countdata as countdata.rds, quit R session and reopened
con <- gzfile('/data/count_all.rds')
countall <- readRDS(con)
FC <- countall[,c(2,7,8)]
library(data.table)
FC2 <- dcast(FC,SEQ ~ CLEANNAME, value.var = 'logFC', fun.aggregate = mean)
mat <- as.matrix(FC2[,c(2:399)])
library(limma)
mat2 <- normalizeQuantiles(mat)
mat[1:5,1:5]
mat2[1:5,1:5]
FC2[,c(2:399)] <- mat2

con <- gzfile('/data/ShrnaGeneMap.RDS','rb')
genemap <- readRDS(con)
# some elements of GENESYMBOLS column have multiple gene names separated by commas; 
# this created a problem later on, when running RSA.R script, so...
genemap$GENESYMBOLS <- gsub(',', '/', genemap$GENESYMBOLS )

joined <- merge(x=genemap,y=FC2,by="SEQ", all.x=T)
saveRDS(joined,'/data/qn_all.RDS')
```
Split the data into csv tables, one per cell line. The RSA.R script will be run on each of them.
```{r}

for (i in c(4:401)) {
  cell <- colnames(joined)[i]
  tab <-  cbind(joined$GENESYMBOLS, joined$SEQ, cell, joined[,i])
  colnames(tab) <- c("Gene_ID", "Well_ID", "Cell", "Score")
  write.table(tab, paste0('/data/RSA/all/',cell,'.csv'), quote=F, eol='\r\n', sep=',', row.names=F)
}

```

The RSA.R script was run in bash.

```{bash}
cd /data/lines
vi job.sh
```
This the job.sh file for running RSA.R on each csv file:
```
#!/bin/bash
FILES=/data/lines/*
for f in *.csv
do
 echo "Processing ${f} file..."
 # take action on each file. $f store current file name
 R CMD BATCH --vanilla --slave --args --l=-2 --u=0 --i=${f} --o=${f}.out -b /home/cbruce/RSA.R
done
```
in bash:
```{bash}
chmod +x job.sh
job.sh
# Combine header of one of the output files with the body (no header) of all output files
(head -1 ymb1.csv.out; tail -n +2 -q *.csv.out ) > all.out
#awk -F',' '$4 < -1 && $5< -1.3 && $6==1' all.out > all.hit.out
awk -F',' '$6==1' all.out > all.hit.out
# 60834714 all.out
# 21292539 all.hit.out

```
## RSA analysis on pools
One of the intermediate files saved earlier is used:
```{r}
con <- gzfile('/data/count_all.rds','rb')
countall <- readRDS(con)
countdata <-countall[,1:7]

treat <- factor(substring(colnames(countdata)[3:4],1,5))
treat <- relevel(treat, ref="PLASM")
design <- model.matrix(~treat)
design

library(edgeR)

count_poolA <- subset(countdata, POOL=="poolA")
y.a <- DGEList(counts=count_poolA[,3:4],genes=count_poolA[,c(1,2,5,6,7)])
fit.a <- glmFit(y.a, design, dispersion=0.2, prior.count=12)
fit2.a <- glmLRT(fit.a)
count_poolA <- cbind(count_poolA,fit2.a$table)

count_poolB <- subset(countdata, POOL=="poolB")
y.b <- DGEList(counts=count_poolB[,3:4],genes=count_poolB[,c(1,2,5,6,7)])
fit.b <- glmFit(y.b, design, dispersion=0.2, prior.count=12)
fit2.b <- glmLRT(fit.b)
count_poolB <- cbind(count_poolB,fit2.b$table)

count_BGPD <- subset(countdata, POOL=="BGPD")
y.c <- DGEList(counts=count_BGPD[,3:4],genes=count_BGPD[,c(1,2,5,6,7)])
fit.c <- glmFit(y.c, design, dispersion=0.2, prior.count=12)
fit2.c <- glmLRT(fit.c)
count_BGPD <- cbind(count_BGPD,fit2.c$table)

# delete unneeded objects to free up memory
rm(countall)
rm(list = c('fit.a','fit.b','fit.c','fit2.a','fit2.b','fit2.c','y.a','y.b','y.c'))

#Restart R, due to memory issues
con <- gzfile('/data/count_poolA.rds','rb')
count_poolA <- readRDS(con)
con <- gzfile('/data/count_poolB.rds','rb')
count_poolB <- readRDS(con)
con <- gzfile('/data/count_BGPD.rds','rb')
count_BGPD <- readRDS(con)
                    
FC <- rbind(count_poolA, count_poolB, count_BGPD)
library(data.table) # much much faster than reshape2 package
FC1 <- as.data.table(FC)
rm(list=c('count_poolA', 'count_poolB', 'count_BGPD'))
FC2 <- dcast(FC1,SEQ+CLEANNAME ~ POOL, value.var = 'logFC', fun.aggregate = mean)
rm(FC)
rm(FC1)
head(FC2)
```
Quantile normalize data for poolA, poolB, and BGPD:
```{r}
library(limma)
mat <- as.matrix(FC2[,c(3:5)])
mat2 <- normalizeQuantiles(mat)
FC3 <- as.data.frame(FC2)
rm(FC2)
FC3[,c(3:5)] <- mat2
con <- gzfile('/data/ShrnaGeneMap.RDS','rb')
genemap <- readRDS(con)
# some elements of GENESYMBOLS column have multiple gene names separated by commas; 
# this created a problem later on, when running RSA.R script, so...
genemap$GENESYMBOLS <- gsub(',', '/', genemap$GENESYMBOLS )

joined <- merge(x=FC3,y=genemap,by="SEQ", all.x=T)
qn_BGPD <- joined[,c(1,2,3,6,7)]
qn_poolA <- joined[,c(1,2,4,6,7)]
qn_poolB <- joined[,c(1,2,5,6,7)]
saveRDS(qn_BGPD,'/data/qn_BGPD.rds')
saveRDS(qn_poolA,'/data/qn_poolA.rds')
saveRDS(qn_poolB,'/data/qn_poolB.rds')

poolA <- as.data.table(qn_poolA)
poolA <- dcast(poolA,GENESYMBOLS+SEQ ~ CLEANNAME, value.var = 'poolA', fun.aggregate = mean)
poolA <- as.data.frame(poolA)


poolB <- as.data.table(qn_poolB)
poolB <- dcast(poolB,GENESYMBOLS+SEQ ~ CLEANNAME, value.var = 'poolB', fun.aggregate = mean)
poolB <- as.data.frame(poolB)

poolC <- as.data.table(qn_BGPD)
poolC <- dcast(poolC,GENESYMBOLS+SEQ ~ CLEANNAME, value.var = 'BGPD', fun.aggregate = mean)
poolC <- as.data.frame(poolC)

for (i in c(3:400)) {
 cell <- colnames(poolA)[i]
 tab <- cbind(poolA$GENESYMBOLS,poolA$SEQ, cell, poolA[,i])
 colnames(tab) <- c("Gene_ID","Well_ID", "Cell", "Score")
 write.table(tab, paste0('/data/RSA/poolA/',cell,'.csv'), sep=',', quote=T, eol='\r\n', row.names=F)
}

for (i in c(3:400)) {
 cell <- colnames(poolB)[i]
 tab <- cbind(poolB$GENESYMBOLS,poolB$SEQ, cell, poolB[,i])
 colnames(tab) <- c("Gene_ID","Well_ID", "Cell", "Score")
 write.table(tab, paste0('/data/RSA/poolB/',cell,'.csv'), sep=',', quote=T, eol='\r\n', row.names=F)
}

for (i in c(3:400)) {
 cell <- colnames(poolC)[i]
 tab <- cbind(poolC$GENESYMBOLS,poolC$SEQ, cell, poolC[,i])
 colnames(tab) <- c("Gene_ID","Well_ID", "Cell", "Score")
 write.table(tab, paste0('/data/RSA/BGPD/',cell,'.csv'), sep=',', quote=T, eol='\r\n', row.names=F)
}
                      
```
In Bash, create the following job.sh script using vi: 
```
#!/bin/bash
FILES=/data/lines/*
for f in *.csv
do
 echo "Processing ${f} file..."
 # take action on each file. $f store current file name
 R CMD BATCH --vanilla --slave --args --l=-2 --u=0 --i=${f} --o=${f}.out -b /home/cbruce/RSA.R
done
```
Run the scripts on the *csv files in each directory:
```{bash}
cd data/poolA
vi job.sh
chmod +x job.sh
cp job.sh ../poolB
cp job.sh ../BGMD
job.sh
```
Then, combine the files into one large file named `poolA.out`, etc. The first line of one of the *.out files is used for the header. Lines 2 to the last line of all files are added to this header line, with the `tail -n +2 q` command.
```{bash}
cd ../poolA
(head -1 yd38.csv.out; tail -n +2 -q *.csv.out ) > poolA.out
#awk -F',' '$4 < -1 && $5< -1.3 && $6==1' poolA.out > poolA.hit.out
awk -F',' '$6==1' poolA.out > poolA.hit.out
# wc -l poolA*out
#   6943283 poolA.hit.out
#  20335489 poolA.out

cd ../poolB
(head -1 yd38.csv.out; tail -n +2 -q *.csv.out ) > poolB.out
#awk -F',' '$4 < -1 && $5< -1.3 && $6==1' poolB.out > poolB.hit.out
awk -F',' '$6==1' poolB.out > poolB.hit.out
# wc -l poolB*out
#   6973834 poolB.hit.out
#  20241793 poolB.out

cd ../BGPD
(head -1 yd38.csv.out; tail -n +2 -q *.csv.out ) > BGPD.out
#awk -F',' '$4 < -1 && $5< -1.3 && $6==1' BGPD.out > BGPD.hit.out
awk -F',' '$6==1' BGPD.out > BGPD.hit.out
# wc -l BGPD*out
#   8463483 BGPD.hit.out
#  24194467 BGPD.out

cd ../poolA
rm *csv*
cd ../poolB
rm *csv*
cd ../BGPD
rm *csv*

```
