---
title: Pre-processing DRIVE data
output: html_document
params:
  count_file: "s3://ps-informatics/bioinformatics/data/Project_DRIVE/.Input/DriveCountData.RDS"
  genemap_file: "s3://ps-informatics/bioinformatics/data/Project_DRIVE/.Input/ShrnaGeneMap.RDS"
  source_description: "Novartis project DRIVE pooled shRNA screen"
  data_source: "data"
  date: !r format(Sys.Date())
---
```{r child='analysis_metadata.Rmd'}
```

+ Use rmarkdown parameters for passing variables into the document
```{r setup, echo=FALSE, results='hide'}
source("setup.R")
proj_dir <- opts_knit$get("proj.dir")
opts_knit$set(data.dir = file.path(proj_dir, params$data_source))
data_dir <- opts_knit$get("data.dir")
src_dir <- opts_knit$get("src.dir")
count_file <- params$count_file
genemap_file <- params$genemap_file
#analysis <- basename(tools::file_path_sans_ext(knitr::current_input()))
analysis <- "preproc"
local_count_file <- file.path(data_dir,
                        basename(count_file))
local_genemap_file <- file.path(data_dir,
                        basename(genemap_file))
# create data directory if it doesn't exist
output_dir <- file.path(opts_knit$get("data.dir"),
                        analysis,
                        params$date)
ifelse(!dir.exists(output_dir), 
       dir.create(output_dir,recursive=TRUE), FALSE)
source(file.path(proj_dir,"src","R","setup.R"))
Sys.setenv(Count_file = count_file,
           Genemap_file = genemap_file,
           Data_Dir = data_dir,
           Src_Dir = src_dir,
           Local_Count_File = local_count_file,
           Local_Genemap_File = local_genemap_file,
           Analysis = analysis,
           Proj_Dir = proj_dir,
           Data_Source = params$data_source,
           UUID = opts_knit$get("uuid"), 
           Run_Date = params$date)
```
Input data download from AWS
```{bash download_input_from_s3, echo=FALSE}
cd $Data_Dir
echo $Local_Count_File
# if file exists don't download
if [ ! -f "$Local_Count_File" ]
then
  aws s3 cp --only-show-errors $Count_file .
  echo 'Saved file here:' $Data_Dir
else
  echo 'File exists...'
fi
echo $Local_Genemap_File
# if file exists don't download
if [ ! -f "$Local_Genemap_File" ]
then
  aws s3 cp --only-show-errors $Genemap_file .
  echo 'Saved file here:' $Data_Dir
else
  echo 'File exists...'
fi

```
##Read in RSA data
We impute plasmid count for elements of PLASMID_COUNT column that are NA, with avg_plasmid_count 
(they were made NA because they had low values)
```{r impute_data}
data_dir
setwd(data_dir)
con <- gzfile('DriveCountData.RDS','rb')
countdata <- readRDS(con)
head(countdata,3)
countdata2 <- subset(countdata, !is.na(PLASMID_COUNT), !is.na(SAMPLE_COUNT) )
countdata2 <- DGEList(counts=countdata2[,3:4],genes=countdata2[,c(1,2,5,6,7)])
num_samples <- dim(countdata2)[1]
total_plasmid_count <- countdata2$samples[1,2]
num_samples
total_plasmid_count
avg_plasmid_count <- round(total_plasmid_count / num_samples)
countdata[is.na(countdata[,3]),3] <- avg_plasmid_count
rm(countdata2)
```
Normalize the plasmid and sample counts using the library sizes.
```{r normalize_plasmid_vs_sample_counts}

countdata2 <- DGEList(counts=countdata[,3:4],genes=countdata[,c(1,2,5,6,7)])
countdata2 <- calcNormFactors(countdata2)

treat <- factor(substring(colnames(countdata)[3:4],1,5))
treat <- relevel(treat, ref="PLASM")
design <- model.matrix(~treat)
```
The edgeR negative binomial model is fitted to obtain log fold change (logFC) of counts
between the sample and the plasmid. This is done for each sample-plasmid pair, to obtain
a logFC per shRNA per cell line.
```{r fit_NB_model}
fit <- glmFit(countdata2, design, dispersion=0.2, prior.count=12)
fit2 <- glmLRT(fit)
head(fit2$table)
countdata <- cbind(countdata,fit2$table)
head(countdata)
```
```{r save_output}
saveRDS(countdata,file.path(data_dir,'countdata.rds'))
```
