---
title: "RSA"
author: "John Bruce"
date: "11/28/2017"
output: html_document
params:
  count_file: "s3://ps-informatics/bioinformatics/data/Project_DRIVE/.Input/DriveCountData.RDS"
  genemap_file: "s3://ps-informatics/bioinformatics/data/Project_DRIVE/.Input/ShrnaGeneMap.RDS"
  source_description: "Novartis project DRIVE pooled shRNA screen"
  data: "data"
  date: !r format(Sys.Date())
---
```{r setup0, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Use rmarkdown parameters for passing variables into the document
```{r set_env, echo=FALSE, results='hide'}
source("setup.R")

proj_dir <- opts_knit$get("proj.dir")
opts_knit$set(root.dir = "proj_dir" )
#opts_knit$set(data.dir = file.path(proj_dir, params$data))
#opts_knit$set(src.dir = file.path(proj_dir, params$source))
data_dir <- opts_knit$get("data.dir")
src_dir <- opts_knit$get("src.dir")
count_file <- params$count_file
genemap_file <- params$genemap_file
#analysis <- basename(tools::file_path_sans_ext(knitr::current_input()))
#analysis <- "preproc"
local_count_file <- file.path(data_dir,
                        basename(count_file))
local_genemap_file <- file.path(data_dir,
                        basename(genemap_file))
# create data directory if it doesn't exist
#output_dir <- file.path(opts_knit$get("data.dir"),
#                        analysis,
#                        params$date)
#ifelse(!dir.exists(output_dir), 
#       dir.create(output_dir,recursive=TRUE), FALSE)
#source(file.path(proj_dir,"src","R","setup.R"))
Sys.setenv(Count_file = count_file,
           Genemap_file = genemap_file,
           Data_Dir = data_dir, 
           Src_Dir = src_dir,
           Local_Count_File = local_count_file,
           Local_Genemap_File = local_genemap_file,
           UUID = opts_knit$get("uuid"),
           Proj_Dir = proj_dir)
```
```{r read_data}
con <- gzfile(file.path(data_dir,basename(count_file)),"rb")
countdata <- readRDS(con)
con <- gzfile("ShrnaGeneMap.RDS","rb")
genemap <- readRDS(con)
close(con)
```
Make a 'wide' table using dcast function; join with genemap table
```{r make_wide_table}
FC <- countdata[,c(2,7,8)]
FC2 <- dcast(FC,SEQ ~ CLEANNAME, value.var = 'logFC', fun.aggregate = mean)
mat <- as.matrix(FC2[,c(2:399)])
mat2 <- normalizeQuantiles(mat)
mat[1:5,1:5]
mat2[1:5,1:5]
FC2[,c(2:399)] <- mat2
```
Process genemap table
```{r process_genemap}
head(genemap,3)
# some elements of GENESYMBOLS column have multiple gene names separated by commas; 
# this created a problem later on, when running RSA.R script, so...
genemap$GENESYMBOLS <- gsub(',', '/', genemap$GENESYMBOLS )

joined <- merge(x=FC2,y=genemap,by="SEQ", all.x=T)
head(joined)
```
Split the data into csv tables, one per cell line. The RSA.R script will be run on each of them.
```{r split_data_into_cell_specific_files}
system('mkdir -p temp')
joined[1:5,1:5]
for (i in c(2:399)) {
  cell <- colnames(joined)[i]
  tab <-  cbind(joined$GENESYMBOLS, joined$SEQ, cell, joined[,i])
  colnames(tab) <- c("Gene_ID", "Well_ID", "Cell", "Score")
  write.table(tab, paste0('temp/',cell,'.csv'), quote=F, eol='\r\n', sep=',', row.names=F)
}
opts_knit$set(root.dir = opts_knit$get("src.dir"))
```
Run the RSA.R code (version 1.8) on each of the cell line files. It will be run with the following parameters from a shell script: 

```{r eval=FALSE} 
R CMD BATCH --vanilla --slave --args --l=-2 --u=0 --i=${f} --o=${f}.out -b ../../src/R/RSA.R
```
(The --l=2 means that any log2FC less than -2 is a definite hit, and --u=0 means any log2FC more than 0 is a definite non-hit)

Run the RSA.R script
```{bash RSA_script}
./RSA_job.sh
```

Combine all the cell-specific files into one file
```{bash tidy_up}
cd temp
onefile=$(ls -1 *csv.out | head -1); (head -1 $onefile; tail -n +2 -q *.csv.out ) > all.rsa.out 
mv all.rsa.out ../
cd ..
rm -rf temp
```
Here is what the output looks like:
```{bash all.rsa.out}
head all.rsa.out
```

